---
title: "Regression Models Project"
author: "Uri Smashnov"
date: "Saturday, February 21st, 2015"
output: pdf_document
---

**Executive Summary**
===

Motor Trend magazine has set to evaluate various parameters that impact cars gas usage as measured by miles per gallon (MPG).  
The research was done using *mtcars* dataset freely available with R installation. The research was set to find answers to the
following questions:   
1) Identify set of significant variables in determining MPG   
2) Is an automatic or manual transmission better for MPG  
3) Quantify the MPG difference between automatic and manual transmissions 

### **Findings**
-Weight(wt), Acceleration(qsec) and type of Transmission(am) are found to be significant at 95% confidence setting  
-Manual transmission has been identified as better for MPG than automatic transmission  
-Manual transmission outperforms automatic transmission by 2.94 MPG on average when holding Weight and Acceleration constant  


**The model and the approach**
===
For the purpose of the research I have used backward step wise approach. I have started by including all variables, and every  
step has eliminated one variable with the highest p-value. To confirm the approach, I have included Anova comparison between the models,  
while Pr(>F) value higher than chosen 5% is indicating that model with higher number of variables don't improve the model.
In the analysis of variance  table we can see that Mode 2 (our final model) is optimal between the models reviewed.  

Final Model  
---  
The Final model and coeficients are presented below.  
```{r cache = TRUE, echo=FALSE}
data(mtcars)
cars=mtcars
cars$am=factor(cars$am)
fit10=lm(data=cars,mpg~I(wt-min(cars$wt))+I(qsec-max(cars$qsec))+am)
fit10
#par(mfrow=c(2,2))
#plot(fit10)
```
Coefficients interpretation
---
- Intercept of 31.7 can be viewed as highest MPG for car with Automatic transmission  
- Addition of every 1,000lb to the weight will reduce MPG by 3.917 MPG. Keeping rest of the parameters unchanged    
- Reduction of every second in acceleration will reduce MPG by 1.23 MPG. Keeping rest of the parameters unchanged    
- Car with Manual transmission will have on average 2.94 better MPG, when holding weight and acceleration constant  

**Confidence intervals for coeficients can be obtained from the Table 1 in Appendix section*  


Residuals analysis  
---   
The analysis reveals that residuals deviate from normal distribution, especially toward the tails. It is also evident that  
the spread of residuals is widens as MPG is growing. It indicates that model lacks certain parameters that can explain high MPG


```{r cache = TRUE, echo=FALSE}
par(mfrow=c(2,2))
plot(fit10)
```



\pagebreak  

**Appendix**    
===   

```{r cache = TRUE, echo=FALSE}
library(ggplot2)
library(car)
library(stargazer)
data(mtcars)
cars=mtcars
cars$am=factor(cars$am)
###### Initial model - including all variables #####################
fit1=lm(data=cars,mpg~cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb)
####################################################################
fit2=lm(data=cars,mpg~disp+hp+drat+wt+qsec+vs+am+gear+carb)
fit3=lm(data=cars,mpg~hp+drat+wt+qsec+vs+am+gear+carb)
fit4=lm(data=cars,mpg~hp+drat+wt+qsec+am+gear+carb)
fit5=lm(data=cars,mpg~hp+drat+wt+qsec+am+carb)
fit6=lm(data=cars,mpg~drat+wt+qsec+am+carb)
fit7=lm(data=cars,mpg~wt+qsec+am+carb)
###### Final model - down to three significant variables #################
fit8=lm(data=cars,mpg~wt+qsec+am)
####included for ilustrative purposes to emphisize that adding of qsec and am is beneficial#####  
fit9=lm(data=cars,mpg~wt) 
#######################################################################
#Model for coefficients interpretation
fit10=lm(data=cars,mpg~I(wt-min(cars$wt))+I(qsec-max(cars$qsec))+am)
summary(fit10)
```

```{r cache = TRUE}
anova(fit9,fit8,fit7,fit5,fit3,fit1)
```

\pagebreak 
```{r cache = TRUE,star, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}
stargazer(fit1, fit4, fit8,title="Regression Models results comparison",dep.var.labels="Models",single.row=TRUE,column.labels=c("All-Var","Mid-Var","Final"),omit.stat=c("LL","ser","f"), ci=TRUE, ci.level=0.95)
```

\pagebreak   

Cross plots of significant variables with computed Correlations   
---  

```{r, echo=FALSE}
###############################################################
# Plot with "pairs" for final model and all paramters in the dataset
######################
m_cars4=data.matrix(cars[,c(1,6,7,9)])
m_cars=data.matrix(cars[,])
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y, use="complete.obs"))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}
panel.hist <- function(x, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks
    nB <- length(breaks)
    y <- h$counts
    y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col="white", ...)
}
pairs(m_cars4, upper.panel = panel.cor,
      diag.panel = panel.hist,
      lower.panel = panel.smooth)
```

**Conclusion and more resources**   
===

The model while providing with relatively high R-square lack in its ability to accurately predict MPG performance at the tals of the scale.  
Is very lickly that some crusial parameters are not present in the data set.  
The code to this report can be obrained from Github repository: <https://github.com/us8945/Regression-Models-Feb-2015>
